Stability and Chaos in Multi-agent Reinforcement Learning

### 2020-10-09 aamas2021-formatting-instructions/aamas21-formatting-instructions/sample.tex
minor mods

\subsection{Related Work}

@@

\paragraph{Evolutionary Game Dynamics}
The theory of evolutionary game dynamics \cite{Morgenstern44}
considers game-like settings in which agents must repeatedly interact
with one another. The outcome of this interaction depends on a payoff
matrix; 'strong' strategies which maximise the reward are promoted,
whilst 'weaker' strategies diminish. The \textit{replicator dynamic}
models this behaviour, and allows one to determine whether, after a
number of iterations, the game is likely to converge to some fixed
equilibrium and, if so, the probabilities with which strategies are
played in this equilibrium (see \cite{ShohamMultiagentFoundations}
for an excellent introduction to the replicator dynamics).

Analyses have been performed from the point of view of the replicator
dynamic towards the various types of behaviours that can emerge from
iterated games. Such games are considered imperfect, in that agents
make decisions by attempting to anticipate their opponents' behaviour
based on experience \cite{Galla2011}. \cite{Imhof2005} presents the
observation that, in the game of Prisoner's Dilemma, cyclic behaviour
emerges in which the optimal strategy cycles across defection,
cooperation and tit-for-tat despite defection being the
NE. In \cite{Galla2011}, Galla builds on this observation by showing that, in fact,
such behaviour is highly dependent on the presence of memory loss in
the system. In the absence of memory loss, the learning converges to
the NE as expected.However, as this parameter is increased, the
stability of the NE is removed and other stable points are created.

Taking the notion of parameter dependence even further, \cite{Galla2013}
describes an analysis of two-person iterated games in which the
players learn using \textit{experience weighted attraction} (EWA), a
form of reinforcement learning typically applied in experimental
economics \cite{Camerer2009}. Through numerical simulations, the
authors were able to show that the emergence of chaos and cycles is
dependent on the choice of parameters (see \cite{Strogatz2000} for
details on chaotic and periodic motion). A rigorous theoretical
analysis of the replicator dynamic corresponding to EWA, which applies
the techniques in \cite{Opper1992}, presented a method for
characterising the regions in parameter space in which a game is
likely to converge, exhibit chaos or limit cycles. More recently, the
same authors presented an extended study in \cite{Sanders2018}, in which
this analysis was extended towards a generic $p$-player game, which
showed that chaotic dynamics are more likely to be observed as the
number of players \textit{p} increases. By contrast to the present work, \cite{Barber2012} focuses on the analysis of the EWA algorithm, while we consider the Q-Learning \cite{Barber2012} algorithm.

It is evident, therefore, that in the case of learning on iterated
games, convergence to stable equilibria cannot be taken for granted
and, in fact, is rare. It would therefore be fruitful to bring these
analyses from EGT to better understand Reinforcement Learning from an
AI perspective.

\paragraph{Dynamics of RL}
In \cite{Tuyls2006AnGames}, the authors are able to derive a relation
between Q-Learning (see \cite{Barber2012}) and the replicator
dynamic. In doing so, the authors present an evolutionary model which
accurately predicts the dynamics of Q-Learning. This sprung forward a
vast array of literature which perform similar derivations for various
reinforcement learning algorithms (see \cite{Bloembergen2015} for an
excellent literature review). In particular it should be noted that
many of these studies focus on normal form games, in which the payoff
matrix does not change (i.e., the environment is static). Recent work,
such as \cite{Hennes2008}, extends this framework towards multi-state
games.
%though this is restricted to 2 state games.
Similarly, \cite{Galstyan2013}, extends the work in
\cite{Tuyls2006AnGames} towards games with continuous action
spaces. Most recently, \cite{Hu2019} considers Q-learning in the
mean-field limit, in which the strategies of populations of agents are
considered, rather than individuals.

The relation between RL and the replicator dynamic allows for the
assumption of convergence in reinforcement learning to be lifted and
instead for the emergence of more complex behaviour to be examined. In
particular, this study aims to perform a similar analysis to that in
\cite{Sanders2018} and establish the regions in parameter space in
which Q-Learning converges to stable fixed points and where the
learning is unstable. In doing so, we extend the dynamics considered by Tuyls et al. towards a general $p$-player scenario, rather than only two players, thereby allowing an analysis of how the number of players in a game affects its stability. To the best of our knowledge, no such work
exists in this context.

@@

    As discussed in the Sec. \ref{sec::Intro}, when learning on iterated
    games, player strategies may exhibit much more complex behaviours
    than convergence to a Nash Equilibrium (NE), including convergence
    to a unique equilibrium (though not always to an NE), convergence
    to one of multiple equilibria, limit cycles, and chaos. These
    behaviours are illustrated in
    Fig.\ref{fig::DynamicalBehaviours}. To be able to predict the
    behaviour of a learning algorithm, it should ideally converge to a
    stable equilibrium although it is still possible to study systems
    with multiple equilibria or limit cycles
    \cite{Strogatz2000}. However, it would be difficult to control
    systems whose dynamics are governed by chaos (though research into
    controlling chaos is ongoing and rife with opportunity
    \cite{Fradkov2009}). It would, therefore, be a useful endeavour to
    determine the conditions under which the different sorts of behaviours
    arise.
%
    \begin{figure}[t]
    \centering
    \includegraphics[width = \linewidth]{Figures/DynamicalBehaviours.png}
    \caption{Evaluation of (\label{fig::DynamicalBehaviours} Different types of dynamical behaviour
       displayed
        by learning agents. Here the $x$-axis (resp.~$y$-axis) is the probability with which agent 1 (resp.~agent 2) chooses a given action. (Top Left) Figure drawn from \cite{Tuyls2006AnGames}.
        Convergence
        to a unique fixed point in the upper right corner (1, 1). This fixed point is unique, in
        that all trajectories, regardless of initialisation, will converge to this point. (Top Right) Limit Cycle, the trajectories converge to cyclic behaviour (Bottom)
        Chaotic behaviour, here small deviations in the initial conditions can grow
        exponentially. (Top Right) and (Bottom) drawn from \cite{Sanders2018}.}
\end{figure}

>>>>

\subsection*{Related Work}

@@

\paragraph{Evolutionary Game Dynamics}
The theory of evolutionary game dynamics \cite{Morgenstern44}
considers game-like settings in which agents must repeatedly interact
with one another. The outcome of this interaction depends on a payoff
matrix; 'strong' strategies which maximise the reward are promoted,
whilst 'weaker' strategies diminish. The \textit{replicator dynamic}
models this behaviour, and allows one to determine whether, after a
number of iterations, the game is likely to converge to some fixed
equilibrium and, if so, the probabilities with which strategies are
played in this equilibrium \cite{ShohamMultiagentFoundations}.
%for an excellent introduction to the replicator dynamics).

Analyses have been performed from the perspective of the replicator
dynamic w.r.t.~the various types of behaviours that can emerge from
iterated games. Such games are considered imperfect, in that agents
make decisions by attempting to anticipate their opponents' behaviour
based on experience \cite{Galla2011}. \cite{Imhof2005} presents the
observation that, in the Prisoner's Dilemma, cyclic behaviour
emerges in which the optimal strategy cycles across defection,
cooperation and tit-for-tat despite defection being the
NE. In \cite{Galla2011}, Galla builds on this observation by showing that, in fact,
such behaviour is highly dependent on the presence of memory loss in
the system. In the absence of memory loss, the learning converges to
the NE as expected.
%However, as this parameter is increased, the stability of the NE is
%removed and other stable points are created.

Taking the notion of parameter dependence further, \cite{Galla2013}
describes an analysis of two-player iterated games in which the
players learn using \textit{experience weighted attraction} (EWA), a
form of reinforcement learning typically applied in experimental
economics \cite{Camerer2009}. Through numerical simulations, the
authors are able to show that the emergence of chaos and cycles is
dependent on the choice of parameters.
%(see \cite{Strogatz2000} for
%details on chaotic and periodic motion).
A rigorous theoretical analysis of the replicator dynamic
corresponding to EWA, which applies the techniques in
\cite{Opper1992}, results in a method for characterising the regions
in parameter space in which a game is likely to converge, exhibit
chaos or limit cycles. More recently, the same authors presented an
extended study in \cite{Sanders2018}, in which the same analysis is
applied to generic $p$-player games, which showed that chaotic
dynamics are more likely to be observed as the number $p$ of players
increases. Differently from this line, we here consider 
%By contrast to the present work, \cite{Barber2012} focuses
%on the analysis of the EWA algorithm, while we consider
 Q-Learning \cite{Barber2012}, which is one of the most popular
 algorithms in RL.

It is evident, therefore, that in the case of learning on iterated
games, convergence to stable equilibria cannot be taken for granted
and, in fact, is rare. It would therefore be fruitful to bring these
analyses from EGT to better understand reinforcement learning from an
AI perspective.

\paragraph{Dynamics of RL}
In \cite{Tuyls2006AnGames}, the authors are able to derive a relation
between Q-Learning
%(see \cite{Barber2012})
and the replicator
dynamic. In doing so, the authors present an evolutionary model which
accurately predicts the dynamics of Q-Learning. This sprung forward a
vast array of literature which performs similar derivations for various
RL algorithms \cite{Bloembergen2015}.
%for an
%excellent literature review).
In particular,
%it should be noted that
many of these studies focus on normal form games, in which the payoff
matrix does not change (i.e., the environment is static). Recent work,
such as \cite{Hennes2008}, extends this framework towards multi-state
games.
%though this is restricted to 2 state games.
Similarly, \cite{Galstyan2013} extends the work in
\cite{Tuyls2006AnGames} towards games with continuous action
spaces. More recently, \cite{Hu2019} considers Q-learning in the
mean-field limit, in which the strategies of populations of agents are
considered, rather than individuals.

The relation between reinforcement learning and the replicator dynamic
allows for the assumption of convergence in RL to be lifted and
instead for the emergence of more complex behaviours to be
examined. In particular, this study aims to perform an analysis
similarly to \cite{Sanders2018} and establish the regions in parameter
space in which Q-Learning converges to stable fixed points.
%and where the learning is unstable.
In doing so, we extend the dynamics considered by Tuyls et al.~to
a general $p$-player setting, rather than two players only, thereby
allowing an analysis of how the number of players in a game affects
its stability.
%To the best of our knowledge, this is the first work exists in
%this context.

@@

    As discussed in the Sec. \ref{sec::Intro}, when learning on iterated
    games, player strategies may exhibit much more complex behaviours
    than convergence to a Nash Equilibrium (NE), including convergence
    to a unique equilibrium (though not always to an NE), convergence
    to one of multiple equilibria, limit cycles, and chaos. These
    behaviours are illustrated in
    Fig.\ref{fig::DynamicalBehaviours}. To be able to predict the
    behaviour of a learning algorithm, it should ideally converge to a
    stable equilibrium although it is still possible to study systems
    with multiple equilibria or limit cycles
    \cite{Strogatz2000}. However, it would be difficult to control
    systems whose dynamics are governed by chaos (though research into
    controlling chaos is ongoing and rife with opportunity
    \cite{Fradkov2009}). It would, therefore, be a useful endeavour to
    determine the conditions under which the different sorts of behaviours
    arise.
%
    \begin{figure}[t]
    \centering
    \includegraphics[width = .85\linewidth]{Figures/DynamicalBehaviours.png}
    \caption{Evaluation of (\label{fig::DynamicalBehaviours} Different types of dynamical behaviour
       displayed
        by learning agents. Here the $x$-axis (resp.~$y$-axis) is the probability with which agent 1 (resp.~agent 2) chooses a given action. (Top Left) Figure drawn from \cite{Tuyls2006AnGames}.
        Convergence
        to a unique fixed point in the upper right corner (1, 1). This fixed point is unique, in
        that all trajectories, regardless of initialisation, will converge to this point. (Top Right) Limit Cycle, the trajectories converge to cyclic behaviour (Bottom)
        Chaotic behaviour, here small deviations in the initial conditions can grow
        exponentially. (Top Right) and (Bottom) drawn from \cite{Sanders2018}.}
\end{figure}
